---
title: "SDS 192 Mini-Project 2"
subtitle: "SDS 192"
author: 
- name: Aviva Jotzke
  affiliation: Smith College
- name: Gigi Ochs
  affiliation: Smith College
date: "`r format(Sys.Date(), '%B %e, %Y')`"
editor: 
  mode: source
output: 
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
execute:
  echo: false
  warning: false
  error: false
---

```{r setup}
library(tidyverse)
library(fec16)
```


In the 2016 election when Donald Trump was running against Hillary Clinton, how secure were incumbent seats for both Democrats and Republicans? We ask this question with the knowledge that Trump ran as the change candidate to Clinton's establishment platform. To what extent was each party craving change from the norm?
```{r analysis-1}
# Group member 1 work here!
house_wnames <- results_house|>
  left_join(candidates, by = c("cand_id"))

house_filtered_r <- house_wnames|>
  select(state, cand_name, incumbent, party, won, general_percent, general_votes, district_id, cand_id)|>
  filter(party == "DEM" | party == "REP")
```

```{r}
# Changing variable names from True/False to terms to make the visualizations easier to read

house_filtered_r$incumb_str <- as.factor( ifelse(house_filtered_r$incumbent == "TRUE", "Incumbent", "Non-Incumbent"))

house_filtered_r$won_str <- as.factor( ifelse(house_filtered_r$won == "TRUE", "Won", "Lost"))

house_filtered_r$party <-
  recode(house_filtered_r$party,
         "DEM" = "Democrat",
         "REP" = "Republican")
```

### Figure 1
```{r}
# Group member 1 
house_filtered_r|>
  # group_by(incumb_str)|>
  ggplot(aes(x = won_str, fill = party))+
  geom_bar(position="dodge")+
  facet_wrap(vars(incumb_str))+
  labs(
    title = "Number of Races Won by Incumbents and Non-Incumbents",
    x = "",
    y = "Count",
    fill = "Party",
    caption = "Source: Federal Election Commission (FEC) 2015-16 Election Cycle Data from Package fec16"
  )+
  scale_fill_manual(values = c("Blue", "Red"))
```

```{r, include=FALSE}
house_filtered_r |>
  filter(won == "TRUE")|>
  group_by(party)|>
  summarize(n = n())
```

```{r analysis-2}
# Group member 2 work here
#This is my join
candidates_joined <- results_house |>
  left_join(candidates,
            by = c("cand_id"))

```

```{r}
#These are my filtered results
house_filteredd <- candidates_joined |>
  select(state, district_id, cand_id, incumbent, party, general_votes, general_percent, won, cand_name, cand_election_yr)|>
  filter(won == "TRUE") |>
  filter(party == "DEM" | party == "REP")
```

```{r}
house_filteredd$incumbent_str <- as_factor(ifelse(house_filteredd$incumbent == "TRUE", "Incumbent", "Non-Incumbent"))
```

### Figure 2

```{r}
#This is my data visualization
  ggplot(
    data = house_filteredd,
    mapping = aes(
         x = general_percent,
         fill = party
  )) +
  geom_histogram(binwidth = 0.1,) +
  labs(title = "Percent of Vote Attained by Winning Incumbents and Non-Incumbents",
       x = "Percent",
       y = "Number of Winning Candidates") +
facet_wrap(vars(party, incumbent_str)) +
scale_fill_manual(values = c("Blue", "Red")) 

```

# Blog post

The dataset used in this project is Fec16, with data taken from the Federal Election Commission (FEC) website. As an agency, the FEC administers federal campaign finance law and moderates financing of the U.S. House, Senate, Presidency, and Vice Presidency campaigns. The Fec16 data set contains specific data relating to candidates, committees, individual and committee contributions, and results for the 2015-2016 election, along with other financial information. While approaching this analysis, our group wondered how secure the incumbent seats were for both Democrats and Republicans in various levels of government when Donald Trump was running against Hillary Clinton. We asked this question with the knowledge that Trump ran as the change candidate to Clinton's establishment platform. We also wanted to understand to what extent each party was craving change from the norm at the time. 

From our first figure, the biggest takeaway is that, in 2016, incumbents from both parties tended to win at higher rates than non-incumbents. It is commonly accepted that incumbents have a political advantage when it comes to Congressional elections and this held true even in 2016 when the country was craving change at the presidential level. Notably, Republican incumbents won at a higher rate than Democrat incumbents though Republican and Democrat non-incumbents won at similar rates. A very small number of incumbents lost their races in both parties, and more Republican non-incumbents ran unsuccessful campaigns than Democrat non-incumbents. From our second figure, the most interesting information is in the incumbent histograms. We see that Republican wins are clustered much closer to 50% than Democrat incumbent wins. This indicates that Republicans tend to more narrowly win their elections while Democrats have much more comfortable leads. This could be a symptom of gerrymandering where partisan groups have redrawn district lines to give their favored party (in this case Republicans) a slight advantage in many districts and confine the other party’s supporters to a single district (leading to large Democrat victories in that district and narrow Democrat losses in others).
	
On the whole, this analysis indicates that Americans were not craving Congressional change to the same extent they were craving change in the presidency. We did not observe a huge wave of incumbent losses nor a huge wave of non-incumbent wins. It is possible, however, that constituents with Republican incumbents were less happy with their representative as Republican incumbent wins tended to be won with a closer margin than Democrat incumbent wins; this may be an effect of gerrymandering. Although we did not notice any ethical concerns in this specific data set, joining data increases the chances of using data not for its original purpose. While the data joined in Fec16 was used to combine similar personal information about specific candidates, the “join” feature could potentially combine data sets that have different purposes and create the appearance that the data sets have more in common than they actually do. 
 
Note: We filtered the data to only include Democrats and Republicans. While this action removes 64 elections from the data set, we believe the data still fits the purpose and scope of our analysis. 

