---
title: "SDS 192 Mini-Project 1"
subtitle: "SDS 192"
author: 
- name: "Aviva Jotzke"
  affiliation: Smith College
- name: "Gigi Ochs"
  affiliation: Smith College
date: "`r format(Sys.Date(), '%B %e, %Y')`"
editor: 
  mode: source
output: 
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
    contained: TRUE
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: FALSE
#| warning: FALSE
# library packages and set up API key

library(tidyverse)
library(rscorecard)
library(RColorBrewer)

sc_key("KvJj3Uee1GY17cccTy36ZcRiq2VbPPeQkNtUZGWl")
```

```{r import, warning=FALSE}
#| echo: FALSE
#| warning: FALSE
# import and clean data 

scorecard <- sc_init() |>

  sc_year(2022) |>                 #Note how we are looking at only 2022 data here!

  sc_filter(stabbr == "MA") |>     #Note how we are looking at only Massachusetts data here!
  #The line below shows variable selection (there are lots of variables)
  sc_select(unitid, instnm, city, highdeg, control, ugds, adm_rate, costt4_a, costt4_p, pcip27, pctfloan, admcon7, wdraw_orig_yr2_rt, cdr3, locale) |>
  sc_get()

# we want instnm, city, highdeg, adm_rate, costt4_a
```

```{r recode}
#  Here's an example of how to recode the control variable

# We are renaming the column control_text from control
scorecard$control_text <-
  #The recode function does the work. It calls 1 "Public", 2 " Private nonprofit",etc. 
  recode(
    scorecard$control, 
    "1" = "Public", 
    "2" = "Private nonprofit", 
    "3" = "Private for-profit",
    .default = NA_character_
  )
```

```{r}
#Recoding `highdeg` 

scorecard$highdeg_text <-
  recode(
    scorecard$highdeg, 
    "0" = "Non-degree-granting",
    "1" = "Certificate degree",
    "2" = "Associate degree",
    "3" = "Bachelor's degree",
    "4" = "Graduate degree",
    .default = NA_character_
  )
```

```{r}
# aggregate cost data

scorecard <- scorecard |>
  mutate (costt4 = costt4_a + costt4_p)
```

```{r}
# Ordering variable `highdeg_text` by degree value. 

scorecard$highdeg_text <- factor(scorecard$highdeg_text, 
                                 levels = c("Non-degree-granting", 
                                            "Certificate degree", 
                                            "Associate degree", 
                                            "Bachelor's degree", 
                                            "Graduate degree"))
                              
```

```{r}
# Recoding `locale' variable rural, town, suburb, city.

scorecard$locale_text <- 
  recode(
    scorecard$locale,
    "11" = "City",
    "12" = "City",
    "13" = "City",
    "21" = "Suburb",
    "22" = "Suburb",
    "23" = "Suburb",
    "31" = "Town",
    "32" = "Town",
    "33" = "Town",
    "41" = "Rural",
    "42" = "Rural",
    "43" = "Rural"
  )
```

```{r}
# Reording `locale_text` 

scorecard$locale_text <- factor(scorecard$locale_text, 
                                 levels = c("Rural", 
                                            "Town", 
                                            "Suburb", 
                                            "City"))
```


```{r}
# Combining cost variables

scorecard <- scorecard |> 
  mutate(costt4 = costt4_a + costt4_p)

scorecard_filtered <- scorecard |>
  filter(costt4 > 0)
```


# Cost of Massachusetts Colleges (2022)
### Figure 1
```{r plot_1}
# Scatter plot showing the amount of colleges with a certain cost
ggplot(
  data = scorecard_filtered,
  mapping = aes( 
    x = costt4,
    )
) +
  geom_histogram() +
  labs(title = "2022 Cost of Massachusetts Colleges", 
       x = "Cost",
       y = "Count",
       caption = "Author: Gigi Ochs",
       subtitle = "Data Source: U.S. Department of Education, College Scorecard")+
theme_minimal()
```

# Cost of Attendance for MA Colleges by their Locale Type (2022)

### Figure 2
```{r plot_2}
# Scatter plot showing the relationship between `costt4` and `locale_text`and colored by `highdeg_text`

scorecard_filtered |> ggplot(
  aes(x=locale_text, 
      y=costt4, 
      color = highdeg_text))+
  geom_jitter(width= 0.1, size=2)+
  labs(
    title= "Cost of Attendance for MA Colleges in 2022 by their Locale Type",
    x = "Locale Type",
    y = "Average Yearly Cost of Attendance",
    color = "Highest Degree Offered",
    caption = "Author: Aviva Jotzke",
    subtitle = "Data Source: U.S. Department of Education, College Scorecard")+
  scale_color_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3"))
```
 
# Blog post

The data set we analyzed was from the College Scorecard project, which is designed to help families and students familiarize themselves with college costs and outcomes. [College Scorecard](https://collegescorecard.ed.gov/assets/InstitutionDataDocumentation.pdf) provides data about colleges as a whole and files with data on specific study fields within colleges. The College Scorecard obtains its data through federal reporting from institutions, tax information, and data on financial aid. We analyze the relationship between the cost of attending an institution relative to the highest degree one can attain and the location of that institution. As a note, we chose to filter out observations where the average yearly cost of attendance was \$0. While it is possible that these colleges offer free tuition, it became apparent–through looking at a sample of colleges–that several schools that reported an average yearly tuition of \$0 did in fact have tuition and likely did not report it. For this reason, we treat these schools as missing data and remove them from our analysis. 

In our first data visualization, we examine the distribution of average yearly cost for colleges in Massachusetts in 2022. The average cost per year is spread between \$0 per year and up to as much as \$80,000 per year. This indicates there is a wide range in the potential cost of education in Massachusetts. Looking deeper, our second data visualization looks at the average yearly cost of attendance for colleges in MA based on the type of locale the school is in (rural, town, suburb, or city), and the highest degree that institution offers. Immediate conclusions from this data set is that most colleges in MA are in suburbs and cities, and colleges that offer higher degrees (graduate and bachelors degrees) tend to be more expensive. 

Overall, this data indicates that, while Massachusetts has a wide range of costs for education, some degree types and institution types tend to be more expensive. On the whole, schools that offer graduate degrees tend to be more expensive than those that offer bachelor's degrees which tend to be more expensive than those that offer associates and certificate degrees. One ethical concern we note is the lack of information about financial aid offerings/merit scholarships offered by these institutions. This data aggregates the financial cost for all students while actual cost likely varies widely between those with different income levels. To make this data more representative and useful to families and students, future iterations should add granularity to cost variables to include levels of financial aid and merit. The College Scorecard dataset is beneficial because of its transparency and clear image of each school’s statistics, which can aid families and students through the college process. 

